The effect of climate forcing on numerical simulations of the Cordilleran ice sheet at the Last Glacial Maximum

### 2014-03-26 reviews/disc-ac1.tex
Add corrections from Irina

Thank you very much for this supportive review. Your comments raise valid points about presentation and interpretation of the results. Please find our response to them below.

@@

In fact, no. In Figs.~8 and~9 of the discussion paper, temperature climatologies are compared without lapse-rate correction. Due to differences in resolution of surface topography between climatological datasets used, this caused appearance of numerous topographic artefacts on paper Fig.~9. Following your comment, we produced updated figures to include a lapse-rate correction of 6\,{$^\circ$}C\,km$^{-1}$, as applied in the simulations.

For paper Fig.~9 (temperature difference maps, Fig.~AC1), lapse-rate correction results in smoother temperature difference maps, where effective temperature discrepancies are no longer overshadowed by local topographic anomalies (Fig.~AC2). Because temperature differences in this new map are much reduced, we adjusted colour scaling to emphasize smaller values (Fig.~AC3).

For paper Fig.~8 (temperature density maps), we now use bilinearly-interpolated temperature maps, rather than the original data (as presented in Fig.~2--4). This results in smoother density plots than those presented in the discussion paper (Fig.~AC4). In a second step, we apply a lapse-rate correction to project all reanalysis data onto the WorldClim (higher resolution) topography (Fig.~AC5). It now becomes clear that much of the discrepancies observed were due to lapse-rate effects.

For consistency, paper Fig.~10 (precipitation density maps) was also updated using bilinearly-interpolated data (this comment's Fig.~AC6). Note that for all density maps, colour maps were changed to allow for a higher level of detail.

@@

A short discussion of these effects is given in Sect.~5.3, in the context of comparison of simulation results to the geomorphological last glacial maximum ice margin. However you raise here an issue that we had not consider, regarding potential ice sheet evolution to more similar states when accounting for precipitation changes. In the East more particularly, where in most of our simulations, the ice margin do not attain a steady configuration, its position after 10\,kyr is largely determined by its rate of advance during the simulation length, which in turn depends on the amount of precipitation received. Therefore, precipitation corrections in this region could significantly slow down the ice margin advance, resulting in more similar configuration after 10\,kyr. We will include a discussion of this effect in Sect.~5.5.

>>>>

Thank you very much for reading our manuscript and writing this supportive review. Your comments raise valid points about presentation and interpretation of the results. Please find our response below, in which we discuss these points with support of updated figures.

@@

In fact, no. In Figs.~8 and~9 of the discussion paper, temperature climatologies are compared without lapse-rate correction. Due to differences in resolution of surface topography between climatological datasets used, this caused appearance of numerous topographic artefacts on paper Fig.~9. Following your comment, we produced updated figures that include a lapse-rate correction of 6\,{$^\circ$}C\,km$^{-1}$, as applied in the simulations.

The implementation of lapse-rate correction in Fig.~9 (temperature difference maps, Fig.~AC1)  results in smoother temperature difference maps, where effective temperature discrepancies are no longer overshadowed by local topographic anomalies (Fig.~AC2). Because temperature differences in this new map are much reduced, we adjusted colour scaling to emphasize smaller values (Fig.~AC3).

Also, the new Fig.~8 (temperature density maps) now use bilinearly-interpolated temperature maps, rather than the original data (as presented in Fig.~2--4). This results in smoother density plots than those presented in the discussion paper (Fig.~AC4). In a second step, we apply a lapse-rate correction to project all reanalysis data onto the WorldClim (higher resolution) topography (Fig.~AC5). It now becomes clear that much of the discrepancies observed were due to lapse-rate effects.

For consistency, Fig.~10 (precipitation density maps) was also updated using bilinearly-interpolated data (this comment's Fig.~AC6). Note that for all density maps, colour scales were changed to allow for a higher level of detail.

@@

A short discussion of these effects is given in Sect.~5.3, in the context of comparison of simulation results to the geomorphological last glacial maximum ice margin. However you raise here a valid concern regarding potential ice sheet evolution to more similar states when accounting for precipitation changes, not considered in the current version of our manuscript. In the East more particularly, where in most of our simulations, the ice margin does not attain a steady-state configuration, its position after 10\,kyr is largely determined by its rate of advance during the simulation length, which in turn depends on the amount of precipitation received. We agree that precipitation corrections in this region could slow down the ice margin advance, resulting in more similar configuration after 10\,kyr. We will include a discussion of this effect in Sect.~5.5.

### 2014-03-26 reviews/disc-ac2.tex
Add custom DOI command that works with TCD
Add corrections from Irina

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}
\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{natbib}
\usepackage{doi}

@@

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.34,0.59,0.82}
%\providecommand{\doi}[1]{doi:\discretionary{}{}{}#1}

@@

% ----------------------------------------------------------------------
\def\referee#1{\bigskip\textcolor{journalname}{\textit{#1}}}

@@

@@

Thank you for your comment. We apologize for the lack of detail on these issues in the discussion paper, which possibly lead to confusion from your side.

@@

where $\nu=1\times10^{21}$\,Pa\,s is mantle viscosity, $\rho_r = 3300$\,kg\,m$^{-3}$ is lithosphere density, $D=5\times10^{24}$\,N\,m is the lithosphere's flexural rigidity, and $\sigma_{zz}$ corresponds to the ice load \citet{bueler-etal-2007}. In the left-hand part of this equation, the first term accounts for mantle relaxation, the second for point-wise isostasy, and the third for elastic flexure. Please refer to \citet{bueler-etal-2007} for a definition of the pseudo-differential operator $|\nabla|$. Because mantle viscosity $\nu$ is non-zero, there exist a time lag between ice sheet growth and the isostatic bedrock response.

We would like to emphasize that most of our simulations do not reach a climatic equilibrium. In fact, simulations ran to climatic equilibrium (not shown in the paper) produce excessive ice cover in continental regions regardless of the forcing climatology used. Moreover, geomorphological and palaeoclimatic records give support for a short-lived Cordilleran ice sheet. Because the modelled ice sheet is in constant expansion, and because of the first term in the above equation, bedrock elevation do not come to equilibrium with the ice sheet configuration in our runs.

@@

Finally, we also would like to emphasize that the sliding relation is controlled by initial bedrock elevation, and is not altered by its evolution through the run. Your comments show that we need more details about computation of bedrock deformation and its effects in the manuscript. These will for sure be added in the revised version.

@@

>>>>

\usepackage[T1]{fontenc}
\usepackage[pdftex]{graphicx}
\usepackage[pdftex]{color}
\usepackage[colorlinks,citecolor=blue]{hyperref}
\usepackage{natbib}

@@

\graphicspath{{../figures/}}
\definecolor{journalname}{rgb}{0.34,0.59,0.82}

@@

% ----------------------------------------------------------------------
% Interactive comment text begins
% ----------------------------------------------------------------------

\def\doi#1{doi:\allowbreak\href{http://dx.doi.org/#1}{#1}}
\def\referee#1{\bigskip\textcolor{journalname}{\textit{#1}}}

@@

Thank you very much for reading and positively commenting on our manuscript. We apologize for the lack of detail regarding isostatic issues, which we discuss below.

@@

@@

where $\nu=1\times10^{21}$\,Pa\,s is mantle viscosity, $\rho_r = 3300$\,kg\,m$^{-3}$ is lithosphere density, $D=5\times10^{24}$\,N\,m is the lithosphere's flexural rigidity, and $\sigma_{zz}$ corresponds to the ice load \citet{bueler-etal-2007}. In the left-hand part of this equation, the first term accounts for mantle relaxation, the second for point-wise isostasy, and the third for elastic flexure. Please refer to \citet{bueler-etal-2007} for a definition of the pseudo-differential operator $|\nabla|$. Due to high mantle viscosity $\nu$, there exist a time lag between the ice sheet growth and the isostatic bedrock response.

We would like to emphasize that most of our simulations do not reach a climatic equilibrium. In fact, simulations ran to climatic equilibrium (not shown in the paper) produce excessive ice cover in continental regions regardless of the forcing climatology used. Moreover, geomorphological and palaeoclimatic records give support for a short-lived Cordilleran ice sheet. Because the modelled ice sheet is in constant expansion, and because of the first term in the above equation, bedrock elevation does not come to equilibrium with the ice sheet configuration in our runs.

@@

Through the till friction angle parameter $\phi$, sliding is controlled by bedrock elevation. This control is meant to reflect the presence of lubricating marine sediment at the ice sheet base. Such sediments are expected below current sea level, and, to a lesser extent, in areas that were submerged during former glacial periods, which is reflected in our simulation set-up by a linear increase of $\phi$ between 0 and 200 meters above sea level, where it reaches its maximum value.

We want to clarify here that $\phi$ is computed only once and for the entire simulation length at the beginning of the run, based on modern bedrock elevation. Because the distribution of marine sediments is not likely to change much during a single phase of ice sheet advance and retreat, sliding is not affected by bedrock elevation changes due to isostatic effects in our runs.

In general, it became clear to us through your comments that we need more details about computation of bedrock deformation and its effects in the manuscript. These will surely be added in the revised version.

@@

% ----------------------------------------------------------------------
% Interactive comment text ends
% ----------------------------------------------------------------------
